working title Compatibility system and stigma size are the main predictors of heterospecific pollen effect

### 2019-02-20 Manuscript_draft/Manuscript_draft.Rmd
more edits paragraph 3

Plant traits are crucial to understand heterospecific pollen effect, but here it is important to notice that these traits can be seen from a donor perpective or a recipient one. In first place, when the donor is considered pollen size and pollen allelopathy could be the two main predictors of effect. How pollen size and allepothaty are involved in heterospecific pollen effect remains unclear yet. For example, Herrera xxxx observed that small pollen is more likely to clogg the stigma, here we argue that although this is possible, the quantities of heterospecific pollen in nature are generally low as mentioned above and bigger pollen is generally related with higher pollen tube growth rate, which add another layer of complexity to the understanding of the overall importance of this trait. Moreover, pollen allelopathy is still an unclear field for pollination ecologist. When the pollen recipeint is taking into account style length, stigmatic area, flower morphology, number of ovules and incompatibility system seems to be the main predictors of effect. In the case of style length species with shorter styles will have greater heterospecific pollen effect due to pollen competition is more likely to occur and therefore ovule interference. Interestingly, differences of heterospecific pollen load has been shown to do not be influenced by flowr morphology (restricted or unrestricted), and the main predictor of the heterospecific pollen load was stigma size (REF). Therefore, if bigger stigmas can receive more hp pollen, a greater quantity of hp can be translated in a possible greater negative effect. For species that are self-incompatible have stronger barriers to heterospecific pollen than self-compatible species @ashman2013. Nonetheless, an effect of foreign pollen is a bit obscured by the variability within species, however species that are strong selfers or strong outcrossers have less variablity in mating systems and predictions of effect could be more realistic (see figure 1 from @whitehead2018).

@@

Species with similar traits are more closely related  XXXXXXXXX [I would say no. species closely related usually have similar traits, specally if those are phylogenetially conserved]().  (Refs? Brown and Mitchell (2001) Arceo-Gómez et al. (2016) Tong and Huang (2016) ).  Several studies predict that the impact of HP transfer is likely to be greater for closely related species (Ashman and Arceo-Gómez (2013)).  Few studies however, have focused on the impacts of heterospecific pollen of distantly related species Thomson et al. (1982) Galen and Gregory (1989) Neiland and Wilcock (1999). Yet, most insects and most stigmas have been found to carry multiple species of foreign pollen with little attention to degree of relatedness (Arceo-Gómez and Ashman (2016); Fang and Huang (2013) ; also cite studies from pollen transfer networks here such as... ).

[here you change topic, new paragraph?]() Further, a majority of plant species are generalist and thus receive visits from multiple different pollinators.  Given these are generally the ones that receive greater loads of heterospecific pollen Fang and Huang (2013) and unrelated species are more likely to coexist  with other species due to less niche overlap (Ref),  understanding the role of foreign pollen from distantly related species thus deserves greater attention in understanding coexistence blah blahXXXXX refs.. Notwithstanding, the effect of heterospecific pollen of far and close related species at community level remains to be explored beyond single pairwise interactions.

**Paragraph 5** Introducing our experiment

In this study we investigated how floral reproductive traits and relatedness mediate the impact of HP transfer by asking the following research questions : To what extent do (i) floral reproductive traits and (ii) relatedness, mediate the impacts of heterospecific pollen on seed set.  We do this by creating an artificial co-flowering community with 10 species belonging to three different families with different traits.

@@

Other idea based on the paper of Aizen 2007:

He explains seed set in this way. O total number of ovules, d fraction of ovules that become seed set, b proportion of pollen grains that reach ovules, p number of pollen grains.

$S=dO(1-e^bp)$

$S=dO(1-e^bp(HpEffect))$

but this Hp effect maybe can be divided at the same time in the interaction between recipient-donor with the main traits that drive the effect. IB: maybe just a weighting factor 0-1 depending on trait matching? If effect of hetero is 0, it cancells out the term, if is maximum, it *1

Moreover, this should consider quantity of hp. lINEAR EFFECT? How to model this maybe talk with nacho. IB: In the absence of data you can try linearity, but also a sigmoidal.

IB: Cool! se also Morris et al 2010 Ecology on how to add to AIzen curve, the effects of a second curve describing the negative effect of hetero. Morris apporach is mathematically more robust if you can model the second curve of hp effects.

I also think you should use this for another paper and look into Morales -Castilla TREE paper inferring interactions. I can see a similar idea where you use a set of matrix you can multiply. 
1) A matrix of plant-plant pollinator influence ( a lo carvalehire 2014). This tells you which plants may get exposed to hp pollen from empirical plant pollinator networks
2) you matrix of plant plant ht effect (0-1) or a proxy based on phylogeny of trait similarity. This is also easy to quantify.
3) A vector of sensitive recipient traits.
 to create the probability of ht effects. I really like this ideas to link to community ecology.

>>>>

Plant traits are crucial to understand heterospecific pollen effect but the multifactorial nature of the traits that are involve in the pollen-pistil interaction make difficult to unravel what are the main traits in driving the effect. On the one hand the main predictive traits of effect from the donor perspective are pollen size, pollen allelopathy and pollen apertures (REFS). In Ashamn 2013 small pollen is predicted to cause a greater fitness decrease, although this can be true there are also other possibilities to consider which can obscure a predictive framework like big pollen can clogg small stigmas with fewer pollen grains, bigger stigmas are less likely to be clogged by small pollen grains and bigger pollen can outcompete smaller pollen grains due to faster pollen tube growth rate (REF). On the other hand  the main recipient traits that can drive heterospecific pollen effect are: stigma size, style length, number of ovules, incompatibility system and flower morphology (REFS). In first place, greater stigmatic area is related with greater amount of heterospecific pollen (REF) and therefore possibly a greater negative effect (REF). Species with shorter styles are more likely to be vulnerable to heterospecific pollen effect (REF). For species that are self-incompatible have stronger barriers to heterospecific pollen than self-compatible species @ashman2013. Nonetheless, an effect of foreign pollen is a bit obscured by the variability within species, however species that are strong selfers or strong outcrossers have less variablity in mating systems and predictions of effect could be more realistic (see figure 1 from @whitehead2018). All in all, there is a need to deepen in the understanding of the relative importance of these traits, due to we still do not know what are the main traits in driving the effect with all these possible combinations that can occur in nature.

@@

Species closely related are more likely to have similar traits XXXXXXXXX.  (Refs? Brown and Mitchell (2001) Arceo-Gómez et al. (2016) Tong and Huang (2016) ).  Several studies predict that the impact of HP transfer is likely to be greater for closely related species (Ashman and Arceo-Gómez (2013)).  Few studies however, have focused on the impacts of heterospecific pollen of distantly related species Thomson et al. (1982) Galen and Gregory (1989) Neiland and Wilcock (1999). Yet, most insects and most stigmas have been found to carry multiple species of foreign pollen with little attention to degree of relatedness (Arceo-Gómez and Ashman (2016); Fang and Huang (2013) ; also cite studies from pollen transfer networks here such as... ).

Further, a majority of plant species are generalist and thus receive visits from multiple different pollinators.  Given these are generally the ones that receive greater loads of heterospecific pollen Fang and Huang (2013) and unrelated species are more likely to coexist  with other species due to less niche overlap (Ref),  understanding the role of foreign pollen from distantly related species thus deserves greater attention in understanding coexistence blah blahXXXXX refs.. Notwithstanding, the effect of heterospecific pollen of far and close related species at community level remains to be explored beyond single pairwise interactions. In this study we investigated how floral reproductive traits and relatedness mediate the impact of HP transfer by asking the following research questions : To what extent do (i) floral reproductive traits and (ii) relatedness, mediate the impacts of heterospecific pollen on seed set.  We do this by creating an artificial co-flowering community with 10 species belonging to three different families with different traits.

@@

### 2019-02-20 Manuscript_draft/Manuscript_draft.tex
more edits paragraph 3

Plant traits are crucial to understand heterospecific pollen effect, but
here it is important to notice that these traits can be seen from a
donor perpective or a recipient one. In first place, when the donor is
considered pollen size and pollen allelopathy could be the two main
predictors of effect. How pollen size and allepothaty are involved in
heterospecific pollen effect remains unclear yet. For example, Herrera
xxxx observed that small pollen is more likely to clogg the stigma, here
we argue that although this is possible, the quantities of
heterospecific pollen in nature are generally low as mentioned above and
bigger pollen is generally related with higher pollen tube growth rate,
which add another layer of complexity to the understanding of the
overall importance of this trait. Moreover, pollen allelopathy is still
an unclear field for pollination ecologist. When the pollen recipeint is
taking into account style length, stigmatic area, flower morphology,
number of ovules and incompatibility system seems to be the main
predictors of effect. In the case of style length species with shorter
styles will have greater heterospecific pollen effect due to pollen
competition is more likely to occur and therefore ovule interference.
Interestingly, differences of heterospecific pollen load has been shown
to do not be influenced by flowr morphology (restricted or
unrestricted), and the main predictor of the heterospecific pollen load
was stigma size (REF). Therefore, if bigger stigmas can receive more hp
pollen, a greater quantity of hp can be translated in a possible greater
negative effect. For species that are self-incompatible have stronger
barriers to heterospecific pollen than self-compatible species Ashman
and Arceo-Gómez (2013). Nonetheless, an effect of foreign pollen is a
bit obscured by the variability within species, however species that are
strong selfers or strong outcrossers have less variablity in mating
systems and predictions of effect could be more realistic (see figure 1
from Whitehead et al. (2018)).

@@

Species with similar traits are more closely related XXXXXXXXX \href{}{I
would say no. species closely related usually have similar traits,
specally if those are phylogenetially conserved}. (Refs? Brown and
Mitchell (2001) Arceo-Gómez et al. (2016) Tong and Huang (2016) ).
Several studies predict that the impact of HP transfer is likely to be
greater for closely related species (Ashman and Arceo-Gómez (2013)). Few
studies however, have focused on the impacts of heterospecific pollen of
distantly related species Thomson et al. (1982) Galen and Gregory (1989)
Neiland and Wilcock (1999). Yet, most insects and most stigmas have been
found to carry multiple species of foreign pollen with little attention
to degree of relatedness (Arceo-Gómez and Ashman (2016); Fang and Huang
(2013) ; also cite studies from pollen transfer networks here such
as\ldots{} ).

\href{}{here you change topic, new paragraph?} Further, a majority of
plant species are generalist and thus receive visits from multiple
different pollinators. Given these are generally the ones that receive
greater loads of heterospecific pollen Fang and Huang (2013) and
unrelated species are more likely to coexist with other species due to
less niche overlap (Ref), understanding the role of foreign pollen from
distantly related species thus deserves greater attention in
understanding coexistence blah blahXXXXX refs.. Notwithstanding, the
effect of heterospecific pollen of far and close related species at
community level remains to be explored beyond single pairwise
interactions.

\textbf{Paragraph 5} Introducing our experiment

In this study we investigated how floral reproductive traits and
relatedness mediate the impact of HP transfer by asking the following
research questions : To what extent do (i) floral reproductive traits
and (ii) relatedness, mediate the impacts of heterospecific pollen on
seed set. We do this by creating an artificial co-flowering community
with 10 species belonging to three different families with different
traits.

@@

Other idea based on the paper of Aizen 2007:

He explains seed set in this way. O total number of ovules, d fraction
of ovules that become seed set, b proportion of pollen grains that reach
ovules, p number of pollen grains.

\(S=dO(1-e^bp)\)

\(S=dO(1-e^bp(HpEffect))\)

but this Hp effect maybe can be divided at the same time in the
interaction between recipient-donor with the main traits that drive the
effect. IB: maybe just a weighting factor 0-1 depending on trait
matching? If effect of hetero is 0, it cancells out the term, if is
maximum, it *1

Moreover, this should consider quantity of hp. lINEAR EFFECT? How to
model this maybe talk with nacho. IB: In the absence of data you can try
linearity, but also a sigmoidal.

IB: Cool! se also Morris et al 2010 Ecology on how to add to AIzen
curve, the effects of a second curve describing the negative effect of
hetero. Morris apporach is mathematically more robust if you can model
the second curve of hp effects.

I also think you should use this for another paper and look into Morales
-Castilla TREE paper inferring interactions. I can see a similar idea
where you use a set of matrix you can multiply. 1) A matrix of
plant-plant pollinator influence ( a lo carvalehire 2014). This tells
you which plants may get exposed to hp pollen from empirical plant
pollinator networks 2) you matrix of plant plant ht effect (0-1) or a
proxy based on phylogeny of trait similarity. This is also easy to
quantify. 3) A vector of sensitive recipient traits. to create the
probability of ht effects. I really like this ideas to link to community
ecology.

>>>>

Plant traits are crucial to understand heterospecific pollen effect but
the multifactorial nature of the traits that are involve in the
pollen-pistil interaction make difficult to unravel what are the main
traits in driving the effect. On the one hand the main predictive traits
of effect from the donor perspective are pollen size, pollen allelopathy
and pollen apertures (REFS). In Ashamn 2013 small pollen is predicted to
cause a greater fitness decrease, although this can be true there are
also other possibilities to consider which can obscure a predictive
framework like big pollen can clogg small stigmas with fewer pollen
grains, bigger stigmas are less likely to be clogged by small pollen
grains and bigger pollen can outcompete smaller pollen grains due to
faster pollen tube growth rate (REF). On the other hand the main
recipient traits that can drive heterospecific pollen effect are: stigma
size, style length, number of ovules, incompatibility system and flower
morphology (REFS). In first place, greater stigmatic area is related
with greater amount of heterospecific pollen (REF) and therefore
possibly a greater negative effect (REF). Species with shorter styles
are more likely to be vulnerable to heterospecific pollen effect (REF).
For species that are self-incompatible have stronger barriers to
heterospecific pollen than self-compatible species Ashman and
Arceo-Gómez (2013). Nonetheless, an effect of foreign pollen is a bit
obscured by the variability within species, however species that are
strong selfers or strong outcrossers have less variablity in mating
systems and predictions of effect could be more realistic (see figure 1
from Whitehead et al. (2018)). All in all, there is a need to deepen in
the understanding of the relative importance of these traits, due to we
still do not know what are the main traits in driving the effect with
all these possible combinations that can occur in nature.

@@

Species closely related are more likely to have similar traits
XXXXXXXXX. (Refs? Brown and Mitchell (2001) Arceo-Gómez et al. (2016)
Tong and Huang (2016) ). Several studies predict that the impact of HP
transfer is likely to be greater for closely related species (Ashman and
Arceo-Gómez (2013)). Few studies however, have focused on the impacts of
heterospecific pollen of distantly related species Thomson et al. (1982)
Galen and Gregory (1989) Neiland and Wilcock (1999). Yet, most insects
and most stigmas have been found to carry multiple species of foreign
pollen with little attention to degree of relatedness (Arceo-Gómez and
Ashman (2016); Fang and Huang (2013) ; also cite studies from pollen
transfer networks here such as\ldots{} ).

Further, a majority of plant species are generalist and thus receive
visits from multiple different pollinators. Given these are generally
the ones that receive greater loads of heterospecific pollen Fang and
Huang (2013) and unrelated species are more likely to coexist with other
species due to less niche overlap (Ref), understanding the role of
foreign pollen from distantly related species thus deserves greater
attention in understanding coexistence blah blahXXXXX refs..
Notwithstanding, the effect of heterospecific pollen of far and close
related species at community level remains to be explored beyond single
pairwise interactions. In this study we investigated how floral
reproductive traits and relatedness mediate the impact of HP transfer by
asking the following research questions : To what extent do (i) floral
reproductive traits and (ii) relatedness, mediate the impacts of
heterospecific pollen on seed set. We do this by creating an artificial
co-flowering community with 10 species belonging to three different
families with different traits.

@@
